Software Citation Principles

### 2016-03-09 sc-principles/software-citation-principles.tex

\author{Force11 Software Citation Working Group (Editors: Arfon M. Smith, Daniel S. Katz, Kyle E. Niemeyer)}

@@

We have documented and analyzed a set of use cases related to software citation in \cite{SC-Use-Cases}.
Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case.

@@

Wilson suggests that software authors include a `CITATION' file that documents exactly how the
authors of the software would like to be cited by others~\cite{ssi-citation-files}. While this isn't a formal metadata specification (it's not machine
readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \cite{10.1371/journal.pone.0136631}.} \katznote{added this ref to the introduction - not sure if it is still needed here too}
In a later post on the SSI blog, Jackson gives a good overview of some of the approaches
package authors have taken to automate the generation of citation entities such as bibtex entries~\cite{ssi-how-shalt-i-cite-thee},
and Knepley at al. do similarly~\cite{knepley2013accurately}.

@@

In the principles, we did not define what software should be cited, but how it should be cited.
What software should be cited is the decision of the author(s) of the research work, in the context of community norms and practices, and in most scientific communities, these are currently in flux.
In general, we believe that any software on which a research product directly depends should be cited, if it significantly contributes to the research product.
Again, the specific decision of what is significant must be made by the author(s) of the product.
However, an illustrative example is the use of Microsoft Excel in research.
We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }

@@

Sou-Cheng Choi, NORC at the University of Chicago \& Illinois Institute of Technology

@@

Matthew B. Jones, NCEAS, University of California,  Santa Barbara

@@

Daniel S. Katz, University of Illinois (co-chair)

@@

Frank J. Seinstra, Netherlands eScience Center

>>>>

\author{Force11 Software Citation Working Group (Editors: Arfon M.~Smith, Daniel S.~Katz, Kyle E.~Niemeyer)}

@@

We have documented and analyzed a set of use cases related to software citation in \cite{SC-Use-Cases}.
Table~\ref{tab:use_cases} summarizes these use cases and makes clear what the requirements are for software citation in each case.
\scnote{Perhaps the long caption of Table 2 can be mostly moved here and be replaced with a shorter version there.}

@@

Wilson suggests that software authors include a `CITATION' file that documents exactly how the
authors of the software would like to be cited by others~\cite{ssi-citation-files}. While this isn't a formal metadata specification (it's not machine
readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.\dmnote{There is evidence from databases that people actually tend to follow such instructions if they exist: \cite{10.1371/journal.pone.0136631}.} \katznote{added this ref to the introduction - not sure if it is still needed here too}
In a later post on the SSI blog, Jackson gives a good overview of some of the approaches
package authors have taken to automate the generation of citation entities such as bibtex entries~\cite{ssi-how-shalt-i-cite-thee},
and Knepley at \scnote{$\rightarrow$ et} al. do similarly~\cite{knepley2013accurately}.

@@

In the principles, we did not define what software should be cited, but how it should be cited.
What software should be cited is the decision of the author(s) of the research work, in the context of community norms and practices, and in most scientific communities, these are currently in flux.
In general, we believe that any software on which a research product directly depends should be cited, if it significantly contributes to the research product.
Again, the specific decision of what is significant must be made by the author(s) of the product.
However, an illustrative example is the use of Microsoft Excel in research.
We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
\scnote{\emph{Do not cite standard office software (e.g. Word, Excel) or programming languages.  Provide references only for specialized software. }---Source: \url{https://owl.english.purdue.edu/owl/resource/560/10/}. Likewise,  on what software not to cite,  generally speaking, we should not cite general software for conducting library research (e.g., JSTOR Mobile App), writing research papers (e.g., \LaTeX), scientific presentations (e.g., Powerpoint) or communications (e.g. Skype).}

@@

Sou-Cheng T.~Choi, NORC at the University of Chicago \& Illinois Institute of Technology

@@

Matthew B.~Jones, NCEAS, University of California,  Santa Barbara

@@

Daniel S.~Katz, University of Illinois (co-chair)

@@

Frank J.~Seinstra, Netherlands eScience Center
